---
title: "Data wrangling"
author: "Pegah Palizban"
date: "2025-10-13"
output: html_document
editor: source
---




## Part one - Wrangling

This tutorial will allow you to explore `dplyr` functionality based on the previous lecture. Every question can be answered with a combination of `|>` pipes. You should refrain from using temporary variables and statements outside of the range of the tidyverse.  

The first part does not require joins or pivots.

### Import the [data from the website](https://biostat2.uni.lu/practicals/data/judgments.tsv).


Assign to the name `judgments` and correct the column types where needed.

```{r}
# Write your answer here
library(tidyverse)
judge_url <- "https://biostat2.uni.lu/practicals/data/judgments.tsv"
judgments <- readr::read_tsv(judge_url)

# spec(judgments)

jud2 <- readr::read_delim(judge_url,
             col_types = 
               cols(start_date = col_date(format = "%d/%m/%Y"),
                    end_date = col_date(format =  "%d/%m/%Y"),
                    finished = "f",
                    condition = "f",
                    gender = "f",
                    logbook = "c", 
                    exclude = "f",
                    moral_judgment = "d",
                    moral_judgment_disgust = "d",
                    moral_judgment_non_disgust = "d",
                    .default = col_integer()))

# jud2
```


### Use `glimpse()` to identify columns and column types.
```{r}
glimpse(jud2)
```

What are the dimensions of the data set.
The dimension is 188 * 158

### Exercise 1 

You can combine the following tasks into a single statement.

#### Select columns
Select `gender`, `subject`, `age` and referring to the `STAI` questionnaire.

```{r}
# Write your answer here
select(jud2, "gender", "subject", "age", starts_with("STAI"))
```


#### Select all subjects with `STAI_pre` greater than 65 and `STAI_post` greater than 40

```{r}
# Write your answer here
jud2 |> 
  filter(STAI_pre > 65) |> 
  filter(STAI_post > 40) |> 
  select("subject")
```


#### Sort the observations by `STAI_pre` so that the subject with the highest score is on top.

```{r}
# Write your answer here
jud2 |> 
  arrange(desc("STAI_pre"))
```

#### For better readability, move the subject, STAI_pre and STAI_post columns to the first positions.

```{r}
# Write your answer here
jud2 |> relocate("subject", "STAI_pre", "STAI_post")
```

#### Check if the data set contains duplicated rows

::: {.callout-tip}
### Tip
Compare the number of rows in the raw input table with the number of rows 
in a de-duplicated table containing unique rows only. 
:::

```{r}
# Your answer in code here
jud2 |>
  distinct(start_date, end_date)
```



### Exercise 2

#### Create a new column called `STAI_pre_category` 

It should contain "low", "normal" and "high" entries depending on the `STAI_pre` values:

  * if `STAI_pre` is less than `25` assign "low",
  
  * if `STAI_pre` is over `65` assign "high",
  
  * for all other values assign "normal".

#### Arrange the table by `STAI_pre` in a descending manner.

```{r}
# Write your answer here
jud3 <- jud2 |> 
  arrange(desc(STAI_pre)) |> 
  mutate(STAI_pre_category = case_when(
    STAI_pre < 25 ~ "low",
    STAI_pre > 65 ~ "high",
    .default = "normal"
  )) |> relocate(STAI_pre_category)
```


### Display distinct values in `STAI_pre` and `STAI_pre_category`.

```{r}
# Write your answer here
jud3 |> 
  distinct(STAI_pre, STAI_pre_category)
```


##### Normalize the values in the REI group
Divide all entries in the REI questionnaire by 5, the maximal value.

```{r}
# Write your answer here
jud2 |> 
  relocate(starts_with("REI")) |> 
  mutate(across(contains("REI"), `/`, 5))
```
# Exercise 3

```{r}
library(tidyverse)
judge_url <- "https://biostat2.uni.lu/practicals/data/judgments.tsv"
judgments <- readr::read_tsv(judge_url)

select(judgments, "subject")
```

# Checking if the rows are different based on the columns
```{r}
nrow(distinct(trees, Girth, Height, Volume)) == nrow(trees)
```

# Adding name to the first column of the tibble
```{r}
as_tibble(swiss, rownames = "Provinces")
```


